---
title: "ADA final code"
author: "audrey toler"
date: "11/28/2021"
output: html_document
---

Dr. Johnson and Dylan: after data cleaning, please skip down to model 1 and model 2 to see the analyses relevant to my objectives.

## Open libraries
```{r}
library(tidyverse)
library(haven)
library(tableone)
library(odds.n.ends)
library(car)
library(blorr)
library(readr)
library(Hmisc)
library(table1)
library(DiagrammeR) # for figure 1
library(stargazer)
library(gtsummary)# to create regression table
library(finalfit) # for tbl regression formatting

```


## Bring in the data from 2020 BRFSS data and select variables
```{r}
# open file
brfss20 <- read_xpt(file = "C:/Users/audre/OneDrive/Documents/FALL-2021/advanced_data_analysis/ADA_project/LLCP2020XPT/LLCP2020.XPT")
```

# Codebook provided in link below:

* https://www.cdc.gov/brfss/annual_data/2020/pdf/codebook20_llcp-v2-508.pdf

## Check variable classes and missing data
```{r}
# check classes
class(brfss20$SEXVAR)
class(brfss20$"_IMPRACE")
class(brfss20$MARITAL)
class(brfss20$HLTHPLN1)
class(brfss20$PERSDOC2)
class(brfss20$MENTHLTH)
class(brfss20$CIMEMLOS)
class(brfss20$"_AGE_G")

# check values of each variable
table(brfss20$SEXVAR)
table(brfss20$"_IMPRACE")
table(brfss20$MARITAL)
table(brfss20$HLTHPLN1)
table(brfss20$PERSDOC2)
table(brfss20$MENTHLTH)
table(brfss20$CIMEMLOS)
table(brfss20$"_AGE_G")

# make smaller brfss data set
brfss20.small <- brfss20 %>%
  select(SEXVAR, "_IMPRACE", MARITAL, HLTHPLN1, PERSDOC2, MENTHLTH, CIMEMLOS, "_AGE_G") %>%
  mutate(SEXVAR = as.factor(SEXVAR)) %>%
  mutate("_IMPRACE" = as.factor("_IMPRACE")) %>%
  mutate(MARITAL = as.factor(MARITAL)) %>%
  mutate(HLTHPLN1 = as.factor(HLTHPLN1)) %>%
  mutate(PERSDOC2 = as.factor(PERSDOC2)) %>%
  mutate(CIMEMLOS = as.factor(CIMEMLOS)) %>%
  mutate("_AGE_G" = as.factor("_AGE_G"))

# save new data set to computer using below code if necessary
#write.csv(brfss20.small, "C:/Users/audre/OneDrive/Documents/FALL-2021/advanced_data_analysis/ADA_project/brfss20small.csv")

# check summary of smaller data set
summary(brfss20.small)

# drop values that represent "Don't know/Not sure" or "Refused" (refer to codebook)
# MARITAL, HLTHPLN1, PERSDOC2, MENTHLTH, CIMEMLOS have missing data
# code age groups below 44 as NA and then drop
# need to change variables to factor (except for MENTHLTH)
# need to shift values one level down (i.e., 1 becomes 0, 2 becomes 1, etc.)
# categorize MENTHLTH
```

## Clean data
```{r}
# clean data
# for later: table(original var, new var, useNA=“always”)
scd20 <- brfss20 %>%
  select(SEXVAR, "_IMPRACE", MARITAL, HLTHPLN1, PERSDOC2, MENTHLTH, CIMEMLOS, "_AGE_G") %>%
  mutate(SEXVAR = case_when(SEXVAR %in% c(1) ~ 0,
                            SEXVAR %in% c(2) ~ 1),
         SEXVAR = factor(SEXVAR, 0:1, c("Male", "Female"))) %>%
  rename(race = "_IMPRACE") %>%
  mutate(race = case_when(race %in% c(1) ~ 0,
                          race %in% c(2) ~ 1,
                          race %in% c(3) ~ 2,
                          race %in% c(4) ~ 3,
                          race %in% c(5) ~ 4,
                          race %in% c(6) ~ 5),
         race = factor(race, 0:5, c("White, Non-Hispanic", "Black, Non-Hispanic", "Asian, Non-Hispanic", "American Indian/Alaska Native, Non-Hispanic", "Hispanic", "Other race, Non-Hispanic"))) %>%
  mutate(MARITAL = na_if(x = MARITAL, y = 9)) %>%
  mutate(MARITAL = case_when(MARITAL %in% c(1) ~ 0,
                             MARITAL %in% c(6) ~ 1,
                             MARITAL %in% c(2, 3, 4, 5) ~ 2),
         MARITAL = factor(MARITAL, 0:2, c("Married", "A member of an unmarried couple", "Single"))) %>%
  mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 7)) %>%
  mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 9)) %>%
  mutate(HLTHPLN1 = case_when(HLTHPLN1 %in% c(1) ~ 0,
                              HLTHPLN1 %in% c(2) ~ 1),
         HLTHPLN1 = factor(HLTHPLN1, 0:1, c("Yes", "No"))) %>%
  mutate(PERSDOC2 = na_if(x = PERSDOC2, y = 7)) %>%
  mutate(PERSDOC2 = na_if(x = PERSDOC2, y = 9)) %>%
  mutate(PERSDOC2 = case_when(PERSDOC2 %in% c(1) ~ 0,
                              PERSDOC2 %in% c(2) ~ 1,
                              PERSDOC2 %in% c(3) ~ 2),
         PERSDOC2 = factor(PERSDOC2, 0:2, c("Yes, only one",  "More than one", "No"))) %>%
  mutate(MENTHLTH = na_if(x = MENTHLTH, y = 77)) %>%
  mutate(MENTHLTH = na_if(x = MENTHLTH, y = 99)) %>%
  mutate(CIMEMLOS = na_if(x = CIMEMLOS, y = 7)) %>%
  mutate(CIMEMLOS = na_if(x = CIMEMLOS, y = 9)) %>%
  mutate(CIMEMLOS = case_when(CIMEMLOS %in% c(2) ~ 0, # make 'no' the reference group for outcome
                              CIMEMLOS %in% c(1) ~ 1),
         CIMEMLOS = factor(CIMEMLOS, 0:1, c("No", "Yes"))) %>%
  rename(age = "_AGE_G") %>%
  mutate(age = na_if(x = age, y = 1)) %>%
  mutate(age = na_if(x = age, y = 2)) %>%
  mutate(age = na_if(x = age, y = 3)) %>%
  mutate(age = case_when(age %in% c(4) ~ 0,
                         age %in% c(5) ~ 1,
                         age %in% c(6) ~ 2),
         age = factor(age, 0:2, c("45 to 54", "55 to 64", "65 or older"))) %>%
  na.omit()

# code to export smaller dataset
#write.csv(scd20,"C:/Users/audre/OneDrive/Documents/FALL-2021/advanced_data_analysis/ADA_project/scd20.csv", row.names = FALSE)

# check new data
summary(scd20)

# check variables for correct categorization (run code without na.omit)
#table(brfss20$"_AGE_G", scd20$age)
#table(brfss20$CIMEMLOS, scd20$CIMEMLOS)
```

## Now change MENTHLTH into a categorical variable
```{r}
# check mental health values to see how to divide into a categorical variable
table(scd20$MENTHLTH)

# now, change MENTHLTH into a categorical variable
scd20$mentcat = cut(scd20$MENTHLTH, c(0, 13, 30, 88))

# check to make sure responses as 88 add up to original
table(scd20$mentcat)


# clean mentcat and age variables
scd20 <- scd20 %>%
  mutate(mentcat2 = case_when(mentcat %in% c('(30,88]') ~ 0,
                             mentcat %in% c('(0,13]') ~ 1,
                             mentcat %in% c('(13,30]') ~ 2),
         mentcat2 = factor(mentcat2, 0:2, c('None', '1 to 13 days', '14+ days')))
```

## Descriptives
```{r}
# make a descriptive table of overall frequencies and percentages
desc.table <- CreateTableOne(data = scd20)
print(desc.table,
      showAllLevels = TRUE)

# change labels for table 1
label(scd20$age)<-"Age group"
label(scd20$SEXVAR)<-"Sex"
label(scd20$race)<-"Race"
label(scd20$MARITAL)<-"Marital status"
label(scd20$HLTHPLN1)<-"Health insurance status"
label(scd20$PERSDOC2)<-"Personal doctor status"
label(scd20$CIMEMLOS)<-"SCD status"
label(scd20$mentcat2) <- "Mental health status"

# make a table 1
table1(~age + SEXVAR + race + MARITAL + HLTHPLN1 + PERSDOC2 + CIMEMLOS|mentcat2, overall="Total", footnote='SD = standard deviation', scd20)
```

## Flow chart
```{r}
grViz("digraph flowchart {

      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle, fontsize=10] 
      #notes indicate how many boxes you will have in your diagram. Since I have two sets of exclusions, I will have three nodes.
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      
      # edge definitions with the node IDs are used to indicate how the rectangle boxes flow from each other. This is a simple flow chart but I could create branching instead as below
      tab1 -> tab2
}
      #This set of code provides the text in each rectangle box.
      [1]: 'Records received from BRFSS data set n = 401,958'
      [2]: 'Excluding 336,829 individuals diagnosed not 45 or older, \\n or with missing data on variables used in analysis n = 65,129'
      ")
```

## Logistic models and analyses

### BASIC MODEL (ONE OUTCOME, ONE PREDICTOR)
```{r}
# check levels of CIMEMLOS
levels(scd20$CIMEMLOS)

# logistic model with mentcat2 as a predictor
mentLogit <- glm(CIMEMLOS ~ mentcat2, data = scd20, family = "binomial")
summary(mentLogit)

ORment <- exp(cbind(OR = coef(mentLogit), confint(mentLogit)))
ORment

# calculate ORs and 95% CIs
odds.n.ends(mentLogit)
```

### Test for assumptions of binary logistic regression

#### Linearity

It is not necessary to test for linearity since all variables in the analysis are categorical.

#### Influences



#### Multicollinearity

It is not necessary to test for multicollinearity since there are only two variables in the model.

#### Model fit of basic model
```{r}
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(mentLogit)

#Hosmer lemeshow goodness of fit test: a significant p value indicates a bad fit
blr_test_hosmer_lemeshow(mentLogit)
```

### Model 1

### Multivariate model with CIMEMLOS as the dependent variable and mental health status, sex, race, health plan status, personal doctor status, marital status, and age as independent variables
```{r}
# logistic model
mentplusLogit <- glm(CIMEMLOS ~ mentcat2 + SEXVAR + race + HLTHPLN1 + PERSDOC2 + MARITAL + age, data = scd20, family = "binomial")
summary(mentplusLogit)

#calculate and print ORs and 95% CIs  
ORmentplus<-exp(cbind(OR = coef(mentplusLogit), confint(mentplusLogit))) #calculate ORs and 95% CIs
ORmentplus #print ORs and 95% CIs

odds.n.ends(mentplusLogit)

# make a formatted table for odds ratios
tbl_regression(mentplusLogit, exponentiate = TRUE, include = mentcat2, label = list(mentcat2 ~ "Mental health status"))
```

### Test for assumptions

#### Linearity

Linearity is not applicable since there are no continuous variables in my analysis.

#### Influence
```{r}
#Cook's D plot
plot(mentplusLogit, which=4, id.n=5, col="red") 

#Cook's D cutoff=0.0015 Or you can define use the rules described in the lecture
cutoff <- 0.0015 

#identify observations with a Cook's Distance greater than cutoff
obs_no <- as.data.frame(cooks.distance(mentplusLogit)) %>%
  mutate(obs_no=row_number()) %>%
  filter(`cooks.distance(mentplusLogit)` > cutoff)
```


#### Multicollinearity
```{r}
vif(mentplusLogit)
```

#### Model fit for model 1
```{r}
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(mentplusLogit)

#Hosmer lemeshow goodness of fit test: a significant p value indicates a bad fit
blr_test_hosmer_lemeshow(mentplusLogit)
```

### Model 2

### Multivariate model with CIMEMLOS as the dependent variable and mental health status, sex, race, health plan status, personal doctor status, marital status, and age PLUS interaction terms for sex and age as independent variables
```{r}
# logistic model
mentEMLogit <- glm(CIMEMLOS ~ mentcat2 + race + HLTHPLN1 + PERSDOC2 + MARITAL + mentcat2*age + mentcat2*SEXVAR, data = scd20, family = "binomial")
summary(mentEMLogit)

#calculate and print ORs and 95% CIs  
ORmentEM<-exp(cbind(OR = coef(mentEMLogit), confint(mentEMLogit))) #calculate ORs and 95% CIs
ORmentEM #print ORs and 95% CIs

odds.n.ends(mentEMLogit)

# make a formatted table for odds ratios
tbl_regression(mentEMLogit, exponentiate = TRUE, label = list(mentcat2 ~ "Mental health status"))
```

### Test for assumptions

#### Linearity

Linearity is not applicable since there are no continuous variables in my analysis.

#### Influence
```{r}
#Cook's D plot
plot(mentEMLogit, which=4, id.n=5, col="red") 

#Cook's D cutoff=0.0015 Or you can define use the rules described in the lecture
cutoff <- 0.0015 

#identify observations with a Cook's Distance greater than cutoff
obs_no_2 <- as.data.frame(cooks.distance(mentEMLogit)) %>%
  mutate(obs_no_2=row_number()) %>%
  filter(`cooks.distance(mentEMLogit)` > cutoff)
```


#### Multicollinearity
```{r}
vif(mentEMLogit)
```

#### Model fit for model 2
```{r}
#Various pseudo R squares, log likelihood, deviance, AIC, BIC
blr_model_fit_stats(mentEMLogit)

#Hosmer lemeshow goodness of fit test: a significant p value indicates a bad fit
blr_test_hosmer_lemeshow(mentEMLogit)
```

### Model 3 (not in abstract or presentation)

#### Balance the 1's and 0's of the outcome
```{r}
# create new dataset where CIMEMLOS = 0
#set.seed(1)
#scd20_zero <- sample_n(scd20[which(scd20$CIMEMLOS == 'No'),], size = 5454,)
#scd20_one <- scd20[which(scd20$CIMEMLOS == 'Yes'),]
#scd20_balance <- rbind(scd20_zero, scd20_one)
#table(scd20_balance$CIMEMLOS)
```

```{r}
# run model on new dataset
#balanceLogit <- glm(CIMEMLOS ~ mentcat2 + SEXVAR + race + HLTHPLN1 + PERSDOC2 + MARITAL + age, data = scd20, family = "binomial")
#summary(balanceLogit)

#odds.n.ends(balanceLogit, predProbPlot=TRUE)
```

```{r}
# ALTERNATIVE CLEAN UP CODING
# clean variables of interest and add labels
# brfss20_SCD <- brfss20 %>%
  # select(SEXVAR, "_IMPRACE", MARITAL, HLTHPLN1, PERSDOC2, MENTHLTH, CIMEMLOS, "_AGE_G") %>%
  # mutate(SEXVAR = recode_factor(.x = SEXVAR,
                            # '1' = 'Male',
                            # '2' = 'Female')) %>%
 # rename(race = "_IMPRACE") %>%
 # mutate(race = recode_factor(.x = race,
                              #'1' = 'White, Non-Hispanic',
                              #'2' = 'Black, Non-Hispanic',
                              #'3' = 'Asian, Non-Hispanic',
                              #'4' = 'American Indian/Alaska Native, Non-Hispanic',
                              #'5' = 'Hispanic',
                              #'6' = 'Other race, Non-Hispanic')) %>%
  #mutate(MARITAL = na_if(x = MARITAL, y = 9)) %>%
  #mutate(MARITAL = recode_factor(.x = MARITAL,
                             #'1' = 'Married',
                             #'2' = 'Divorce',
                             #'3' = 'Widowed',
                             #'4' = 'Separated',
                             #'5' = 'Never married',
                             #'6' = 'A member of an unmarried couple')) %>%
  #mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 7)) %>%
  #mutate(HLTHPLN1 = na_if(x = HLTHPLN1, y = 9)) %>%
  #mutate(HLTHPLN1 = recode_factor(.x = HLTHPLN1,
                              #'1' = 'Yes',
                              #'2' = 'No')) %>%
  #mutate(PERSDOC2 = na_if(x = PERSDOC2, y = 7)) %>%
  #mutate(PERSDOC2 = na_if(x = PERSDOC2, y = 9)) %>%
  #mutate(PERSDOC2 = recode_factor(.x = PERSDOC2,
                             #'1' = 'Yes, only one',
                             #'2' = 'More than one',
                             #'3' = 'No')) %>%
  #mutate(MENTHLTH = na_if(x = MENTHLTH, y = 77)) %>%
  #mutate(MENTHLTH = na_if(x = MENTHLTH, y = 99)) %>%
  #mutate(CIMEMLOS = na_if(x = CIMEMLOS, y = 7)) %>%
  #mutate(CIMEMLOS = na_if(x = CIMEMLOS, y = 9)) %>%
  #mutate(CIMEMLOS = recode_factor(.x = CIMEMLOS,
                             # '1' = 'Yes',
                             # '2' = 'No')) %>%
  #rename(age = "_AGE_G") %>%
  #mutate(age = recode_factor(.x = age,
                             # '1' = '18 to 44',
                              #'2' = '18 to 44',
                              #'3' = '18 to 44',
                              #'4' = '45 to 54',
                              #'5' = '55 to 64',
                             # '6' = '65 or older'))

# make a subset of data set
#dataSCD20 <- brfss20_SCD[ which(brfss20_SCD$age=='45 to 54' | brfss20_SCD$age=='55 to 64' | brfss20_SCD$age=='65 or older'), ]

#or you can use the below code
#dataSCD20_2 <- subset(brfss20_SCD, age == '45 to 54' | age == '55 to 64' | age == '65 or older',
                    #select = c(SEXVAR, race, MARITAL, HLTHPLN1, PERSDOC2, MENTHLTH, CIMEMLOS, age))

# check mental health values to see how to divide into a categorical variable
#table(dataSCD20$MENTHLTH)

# now, change MENTHLTH into a categorical variable
#dataSCD20$mentcat = cut(dataSCD20$MENTHLTH, c(0, 13, 30, 88))

# check to make sure responses as 88 add up to original
#table(dataSCD20$mentcat)


# clean mentcat and age variables
#dataSCD20 <- dataSCD20 %>%
  #mutate(mentcat2 = case_when(mentcat %in% c('(30,88]') ~ 0,
                             #mentcat %in% c('(0,13]') ~ 1,
                             #mentcat %in% c('(13,30]') ~ 2),
         #mentcat2 = factor(mentcat2, 0:2, c('None', '1 to 13 days', '14+ days'))) %>%
  #mutate(age = na_if(x = age, y = '18 to 44'))

```

